---
title: "Exploring genome variation using GEMINI framework in R"
author: "Sergey Naumenko, CCM Bioinformatics Training series"
date: "November 30, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
library(knitr)
#knitr::opts_chunk$set(echo = T)
```

## What is database?

- Database is a collection of related dataframes (tables);
- GEMINI uses SQLite database which does not require database server or database engine;
- database slang:
    + dataframe = table
    + column = field
    + select * from [table_name] where [condition];
- Gemini Database Schema
  http://gemini.readthedocs.io/en/latest/content/database_schema.html
- Web Interface to plot database schemas:
  http://ondras.zarovi.cz/sql/demo/?keyword=default

## How to load your variants into the GEMINI database?

- Ask your bioinformatics provider to send results in the GEMINI format;
- Ask CCM to convert vcf2gemini;
- Convert
    + gemini load (linux);
    + vcf2db.py 
- Use bcbio pipeline system – it outputs GEMINI and vcf.

## Bcbio pipelines

- open source, community supported
- developed mainly by Harvard T.Chan Medical School Bioinformatics Core
- validated
- well documented
- Pipelines
```{r pipelines,echo=F}
pipelines=data.frame(c("variant calling","cancer","structural variant calling","RNA-seq","smallRNA-seq","ChIP-seq"))
kable(pipelines)
```
- Resources
    + https://bcbio-nextgen.readthedocs.io/en/latest/
    + https://github.com/chapmanb/bcbio-nextgen
    + https://bcb.io

## Example queries
select gene,transcript from gene_detailed where gene='AGRN';
select variant_id,gene,transcript from variant_impacts where variant_id=2;

## Additional columns not present in GEMINI

- UCSC hyperlink
- OMIM gene description (registration required)
- Orphanet status
- exac pLi score
- exec missense score
- phastcons score
- imprinting status
- information from HGMD database (public version is unusable, pro version is pricy)
- pseudoatosomal gene

## Slide with R Output

```{r cars, echo = TRUE}
kable(summary(cars))
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

